Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-02-22 resex.tex

\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
%floatfix,
]{revtex4-1}

@@

\author{Miriam Fitterer}
 \email{mfittere@fnal.gov}
\author{Giulio Stancari}
\author{Alexander Valishev}
\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA}

@@

For an axially symmetric electron lens, the field at the beam core
vanishes. Effects on the proton core arise from imperfections, with
two main sources: the injection and extraction bends of the HEL
(discussed below in Section~\ref{core:sec:1}), where the electron beam
crosses the proton beam; and distortions of the electron beam profile
during its propagation under magnetic focusing and space charge
(Section~\ref{core:sec:2}). Both sources result in nonlinear
kicks~\cite{hel_bends_stancari, hel_model_polynomial_morozov}. In
continuous operation, these nonlinear kicks are usually much smaller
than the machine nonlinearities. Tolerances on imperfections are
therefore not particularly stringent. The picture changes
significantly in case of pulsed operation. If the electromagnetic
field does not vanish at the proton beam core, noise or resonant kicks
are transferred not only to the halo particles, as intended, but also
to the beam core. Tolerances on the residual fields in this case
become much more stringent. Studies of the effects of the HEL on the
beam core therefore focus on this mode of operation, which is also the
main subject of this paper.

@@

\section{Experimental and simulation setup\label{sec:exp}}
% put picture on top

\begin{table*}
	\caption{\label{tab:md_param}%
		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and~2017~\cite{resexmd2016,resexmd2017}. The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
	}
	\begin{ruledtabular}
		\begin{tabular}{lcc}
			Parameter & Experiment 2016 & Experiment 2017  \\
			\colrule
			beam &\multicolumn{2}{c}{Beam~1} \\
			beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
			single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
			normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
			number of bunches & $12\times4=48$ single bunches & $3\times72=216$ bunches\\
			&  & (+ 1 pilot + 12 nominal) \\\hline
			injection optics, $\beta^*=11$~m & standard optics 2016 & standard optics 2017\\
			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=+19.6~\mathrm{A}$ for MOF circuit and}\\
			& \multicolumn{2}{c}{$I_{\mathrm{MO}}=-19.6~\mathrm{A}$ for MOD circuit (standard 2016 settings)}\\\hline
			working point $(Q_x,Q_y)$ & (64.28,59.31) & (62.27,60.295) \\
			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
			pulsing patterns  &$7^{\mathrm{th}}$~turn H; &$7^{\mathrm{th}}$~turn H,V,H+V; \\
			& $10^{\mathrm{th}}$~turn V; & $8^{\mathrm{th}}$~turn H,V,H+V; \\
			& &  random  H,V,H+V;\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}

\begin{figure*}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		2016 experiment\\
		\includegraphics[width=0.9\linewidth]{bunchfilling_2016.png}	
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		$ $\\
		2017 experiment\\
		\includegraphics[width=0.9\linewidth]{bunchfilling_2017.png}	
	\end{minipage}
	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom). In 2016 a total of 48~bunches was used and in 2017 a total of 216~bunches. Each bunch is indicated with a cyan vertical bar. The bunches were grouped in sub-groups of 4~for 2016 and 6~for 2017 bunches, where each sub-group experienced the same excitation pattern and amplitude. The excitation amplitude is shown in blue and red. In 2016 the excitation was only applied in one plane. In 2017 more bunches could be injected without compromising machine protection making it possible to test all three planes in one fill.}
\end{figure*}
%put picture on top

@@

\bibliography{resex-bibdesk}

>>>>

\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
%floatfix,
longbibliography,
]{revtex4-1}

@@

\author{Miriam Fitterer}
 \email{Miriam.Fitterer@cern.ch}
\author{Giulio Stancari}
\author{Alexander Valishev}
\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA}

@@

For an axially symmetric electron lens, the field at the beam core
vanishes. Effects on the proton core arise from imperfections, with
two main sources: the injection and extraction bends of the HEL
(discussed below in Section~\ref{core:sec:1}), where the electron beam
crosses the proton beam; and distortions of the electron beam profile
during its propagation under magnetic focusing and space charge
(Section~\ref{core:sec:2}). Both sources result in nonlinear
kicks~\cite{hel_bends_stancari,hel_model_polynomial_morozov}. In
continuous operation, these nonlinear kicks are usually much smaller
than the machine nonlinearities. Tolerances on imperfections are
therefore not particularly stringent. The picture changes
significantly in case of pulsed operation. If the electromagnetic
field does not vanish at the proton beam core, noise or resonant kicks
are transferred not only to the halo particles, as intended, but also
to the beam core. Tolerances on the residual fields in this case
become much more stringent. Studies of the effects of the HEL on the
beam core therefore focus on this mode of operation, which is also the
main subject of this paper.

@@

\section{Experimental and simulation setup}
\label{sec:exp}

\begin{table*}
  \caption{ Beam parameters and machine configuration as used in the
    two resonant excitation experiments in 2016
    and~2017~\cite{resexmd2016,resexmd2017}. The plane of the
    excitation is abbreviated with H for horizontal, V for vertical
    and H+V horizontal and vertical at the same time.}
  \label{tab:md_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Parameter & Experiment 2016 & Experiment 2017  \\
      \colrule
      beam &\multicolumn{2}{c}{Beam~1} \\
      beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
      single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
      normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
      $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
      $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
      number of bunches & $12\times4=48$ single bunches & $3\times72=216$ bunches\\
                &  & (+ 1 pilot + 12 nominal) \\\hline
      injection optics, $\beta^*=11$~m & standard optics 2016 & standard optics 2017\\
      Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=+19.6~\mathrm{A}$ for MOF circuit and}\\
                & \multicolumn{2}{c}{$I_{\mathrm{MO}}=-19.6~\mathrm{A}$ for MOD circuit (standard 2016 settings)}\\\hline
      working point $(Q_x,Q_y)$ & (64.28,59.31) & (62.27,60.295) \\
      chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
      pulsing patterns  &$7^{\mathrm{th}}$~turn H; &$7^{\mathrm{th}}$~turn H,V,H+V; \\
                & $10^{\mathrm{th}}$~turn V; & $8^{\mathrm{th}}$~turn H,V,H+V; \\
                & &  random  H,V,H+V;\\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

\begin{figure*}
  \begin{minipage}[t]{1.0\linewidth}
    \centering
    2016 experiment\\
    \includegraphics[width=0.9\linewidth]{bunchfilling_2016.png}	
  \end{minipage}
  \begin{minipage}[t]{1.0\linewidth}
    \centering
    $ $\\
    2017 experiment\\
    \includegraphics[width=0.9\linewidth]{bunchfilling_2017.png}	
  \end{minipage}
  \caption{Filling and excitation scheme for LHC experiments in 2016
    (top) and 2017 (bottom). In 2016 a total of 48~bunches was used
    and in 2017 a total of 216~bunches. Each bunch is indicated with a
    cyan vertical bar. The bunches were grouped in sub-groups of 4~for
    2016 and 6~for 2017 bunches, where each sub-group experienced the
    same excitation pattern and amplitude. The excitation amplitude is
    shown in blue and red. In 2016 the excitation was only applied in
    one plane. In 2017 more bunches could be injected without
    compromising machine protection making it possible to test all
    three planes in one fill.}
  \label{fig:fill}
\end{figure*}

@@

%\bibliography{resex-bibdesk}
\bibliography{resex-mf}
